---
title: "Reproducible Research: Peer Assessment 1"
---

#Loading and preprocessing the data

####Reading data

```{r, include=TRUE}
rawdata <- read.csv('activity.csv')
```


####The data is then aggregated by date 


```{r, include=TRUE}
data <- aggregate(steps ~ date, data=rawdata, sum, na.rm = TRUE)
```

####histogram produced displaying the data. This histogram is devided into 20 buckets, which was arbitrarily chosen.

```{r, include=TRUE}
hist(data$steps, breaks=20, main="Total Steps per Day", xlab="Steps", ylab="Frequency")
```

#What is mean total number of steps taken per day?

```{r, include=TRUE}
rawsteps_mean <- mean(rawdata$steps, na.rm=TRUE)
rawsteps_median <- median(rawdata$steps, na.rm=TRUE)
print(paste("The mean steps per day is: ", rawsteps_mean))
print(paste("The median steps per day is: ", rawsteps_median))
```

#What is the average daily activity pattern?

```{r, include=TRUE}
stepsdata <- aggregate(steps ~ interval, data=rawdata, mean, na.rm=TRUE)
```



```{r, include=TRUE}
plot(stepsdata$interval, stepsdata$steps, type="l", main="Average Steps per Five Minute Interval",
     xlab="Interval No.", ylab="steps")
```

####Calculating maximum steps

```{r, include=TRUE}
maxsteps <- max(stepsdata$steps)

print(paste("The maximum number of steps in a five minute interval was: ", maxsteps))
```

#Imputing missing values

```{r, include=TRUE}
missingdata <- sum(is.na(rawdata$steps))

print(paste("There are", missingdata, "missing data points."))
```


```{r, include=TRUE}

betterdata <- rawdata
betterdata$steps[is.na(betterdata$steps)] <- median(rawdata$steps, na.rm=TRUE)
betterdataday <- aggregate(steps ~ date, data=betterdata, sum, na.rm=TRUE)
hist(betterdataday$steps, breaks=20, main="Total Steps per Day \n Adjusted Data",
     xlab="Steps", ylab="Frequency")
```




```{r, include=TRUE}
bsteps_mean <- mean(betterdata$steps)
bsteps_median <- median(betterdata$steps)
print(paste("The mean is: ", bsteps_mean))

```



#Are there differences in activity patterns between weekdays and weekends?

```{r, include=TRUE}
betterdata$date <- as.Date(betterdata$date)
betterdata$dayname <- weekdays(betterdata$date)
betterdata$weekend <- as.factor(ifelse(betterdata$dayname == "Saturday" |
                                         betterdata$dayname == "Sunday", "weekend", "weekday"))
library(lattice)
plotdata <- aggregate(steps ~ interval + weekend, betterdata, mean)
xyplot(steps ~ interval | factor(weekend), data=plotdata, aspect=1/3, type="l")
```
